What We Eval in the Shadows

### 2021-08-14 paper/main.tex
Grammar (end)

Table~\ref{tab:se-env} shows environment kinds where side-effects happen.
Function environments distinguish between \emph{local}, the caller environment
(offset 0), and \emph{function} (offset >0). \emph{Synthetic} represents
constructed environments. \emph{Object} is used to denote the environments
attached to objects and classes in the S4 and R6 object systems. \emph{Multiple}
denotes cases where side-effects to more than one kind originate from one site.
The table gives the number of call sites of \eval (and the ratio) performing
side-effect in a particular environment kind. The table also gives the number of
functions in which these sites occur (and ratio).

@@

The R language was intended to be extensible. The combination of lazy
evaluation, \k{substitute} and \k{eval} are some of the tools given to
developers to this end. To understand \emph{why} developers use \eval, we
manually inspected top 117 sites which contribute to 90\% of the calls in the
corpus. We present examples from nine classes from this batch. The
classification is not exhaustive; we only study 117 call sites from the
\PkgEvalCallSites call sites in our corpus and many uses of \eval do not fit in
a specific category. The goal here is to demonstrate the wide variety of uses to
which \eval is put to use. This also speaks to the ubiquity of \eval in the R
ecosystem. The code snippets have been simplified for clarity.

@@

Non-standard scoping is an important use case of \eval in R. It enables the
design of functions which require less typing. Since R is used primarily
interactively, this trade-off is justifiable.

@@

The cases where \eval appears to be an overkill include simple expression shapes
such as \k{X}, \k{V}, \k{\$}, \k{<-}, and \k{FUN}. Variable lookup in a remote
environment can be performed by the builtin \k{get} function, a similar case can
be made for assignment and the \k{assign} function. There is a dedicated
\k{do.call} function to perform function calls. However, the unnecessary uses of
\eval are not easy to find and often can only be uncovered by manual analysis.
Dynamic analysis is limited due to coverage issues, and static analysis of R
remains an open problem.

>>>>

Table~\ref{tab:se-env} shows environment kinds where side-effects happen.
Function environments distinguish between \emph{local}, the caller environment
(offset 0), and \emph{function} (offset >0). \emph{Synthetic} represents
constructed environments. \emph{Object} is used to denote the environments
attached to objects and classes in the S4 and R6 object systems. \emph{Multiple}
denotes cases where side-effects to more than one kind originate from one site.
The table gives the number of call sites of \eval (and the ratio) performing
side-effects in a particular environment kind. The table also gives the number of
functions in which these sites occur (and ratio).

@@

The R language was intended to be extensible. The combination of lazy
evaluation, \k{substitute} and \k{eval} are some of the tools given to
developers to this end. To understand \emph{why} developers use \eval, we
manually inspected the top 117 sites which contribute to 90\% of the calls in the
corpus. We present examples from nine classes from this batch. The
classification is not exhaustive; we only study 117 call sites from the
\PkgEvalCallSites call sites in our corpus and many uses of \eval do not fit in
a specific category. The goal here is to demonstrate the wide variety of uses to
which \eval is put to use. This also speaks to the ubiquity of \eval in the R
ecosystem. The code snippets have been simplified for clarity.

@@

Non-standard scoping is an important use case of \eval in R. It enables the
design of functions which require less writing. Since R is used primarily
interactively, this trade-off is justifiable.

@@

The cases where \eval appears to be an overkill include simple expression shapes
such as \k{X}, \k{V}, \k{\$}, \k{<-}, and \k{FUN}. Variable lookup in a remote
environment can be performed by the builtin \k{get} function. A similar case can
be made for assignment and the \k{assign} function. There is a dedicated
\k{do.call} function to perform function calls. However, the unnecessary uses of
\eval are not easy to find and often can only be uncovered by manual analysis.
Dynamic analysis is limited due to coverage issues, and static analysis of R
remains an open problem.
